\mytitle

### 2015-08-03 MSc_diss.tex
Added figures from Neu2008

The first 3-D chemical transport  model simulations examining the role of methyl and ethyl nitrates in tropospheric ozone chemistry have been conducted by \citep{Neu2008}. In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM [Carver et al., 1997] and compared their simulated abundances to observations. To assess  the relative importance of
oceanic alkyl nitrates as a natural source of $NO_x$ in remote regions they evaluated their impact on tropospheric $O_3$ and the oxidative capacity of the atmosphere.

Observations have revealed widespread emissions of alkyl nitrates from the tropical and Southern Oceans. We present the first chemical transport model simulations to examine the global impact of these emissions. Matching observed atmospheric abundances, we derive a total oceanic flux of methyl nitrate (MeONO2) and ethyl nitrate (EtONO2) equivalent to 0.35 Tg of N per year, which contributes as much as 1 DU to the tropospheric ozone column in the Western Pacific and is responsible for about 3\% of the global oxidative capacity of the troposphere.
Atmospheric observations indicate that low molecular weight (C1-C4) alkyl nitrates are ubiquitous over the tropical and Southern Oceans [Atlas et al., 1993; Blake et al., 1999, 2003] and can be a significant component (20–80\%) of the total reactive nitrogen in remote marine air [Jones et al., 1999; Talbot et al., 2000]. Recent ship-borne measurements show supersaturation of alkyl nitrates in tropical surface waters, confirming an oceanic source [Chuck et al., 2002; Moore and Blough, 2002; Dahl et al., 2005]. This work presents the first 3-D chemical transport model (CTM) simulations to examine the role of oceanic alkyl nitrates in global chemistry, using observational constraints.
In this study, we incorporate MeONO2 and EtONO2 in the chemical mechanism of the UCI CTM [Carver et al., 1997] and compare their simulated abundances to observations to derive new estimates of their oceanic fluxes. We assess the relative importance of oceanic alkyl nitrates as a natural source of NOx in remote regions by evaluating their impact on tropospheric ozone (O3) and the oxidative ca- pacity of the atmosphere.
The UCI CTM [Wild and Prather, 2000; Hsu et al.,2005] includes 30 chemical species in the troposphere.
We define a control run with no alkyl nitrate emis-
sions and compare it to two simulations in which we impose a constant, spatially uniform flux of MeONO2 and EtONO2 from: 1) the tropical oceans (10?S–10?N) and 2) the Southern Ocean (south of 45?S).
Results: impact on NOx and O3 chemistry
Oceanic alkyl nitrates cause measurable increases in
both NOx and O3 in marine air. Their largest impact is seen over the equatorial Western Pacific (Figure 3). There is a persistent minimum in both NOx and O3 (Figure 3a) in this region, where the circulation isolates the marine boundary layer, resulting in sustained photochemical loss of O3 and a deep minimum throughout the tropospheric O3 column. At the same time, the circulation acts to accumulate alkyl nitrates in theWestern Pacific (see Figure 2), and they drive up boundary layer NOx abundances by as much as 250\%. This leads to O3 increases of up to 20\% (Figures 3b and 3f), with the largest increases occurring from May to October. From December to March (not shown), the relative increases in NOx and O3 are smaller, but are still concen- trated in the Western Pacific even though the alkyl nitrates are spread more evenly across the Pacific during these months. The geographic pattern of the alkyl nitrate influence persists into the middle and upper troposphere (Figures 3c-3d and 3g-3h), with reduced relative impact because of the larger lightning NOx source. However, the largest zonal mean increases in mixing ratio, 3–8 ppt for NOx and 1 ppb for O3, occur in the tropical upper and middle troposphere, respectively (Figure 4).
Pacific for example, ignore Southern ocean.
The O3 produced by NOx from alkyl nitrates fills in
the minimum in the tropospheric O3 column over the Western Pacific, adding more than 1 DU, with a maximum in October (Figure 3e). The global average increase in tropospheric O3 column varies with season from 0.35 to 0.42 DU.!!!!
WOW!!!
Oceanic alkyl nitrates are necessary to accurately
model NOx in the tropical remote marine environment, since they are responsible for 15–20\% of the O3 in the Western Pacific. The coincident location of the maximum alkyl nitrate impact and the deep minimum in O3 over the Western Pacific is a fascinating consequence of the convergence gence of the trade winds, but will make it difficult to detect an alkyl nitrate signature in O3. [22] At current levels, alkyl nitrates have a modest climate impact. From a simple scaling [Wild et al., 2001], the global increase in tropospheric column O3 will have a radiative forcing of about 16 mW/m2. In steady state, this will be almost entirely offset by the reduced CH4 lifetime, which produces a roughly equal negative radiative forcing. Surprisingly, a comparison to Wild et al. [2001] shows that on a per-molecule basis, alkyl nitrates have the largest impact on O3 and CH4 of any NOx source analyzed to date. This seems to be a consequence of their long lifetime combined with efficient transport to the tropical upper troposphere. With a 2.4\% effect on CH4 lifetime, a 5-fold increase in alkyl nitrate emissions during glacial periods would account for roughly 1/3 of the 300 ppb drop in CH4 seen between interglacials and glacials.

>>>>

The first 3-D chemical transport  model simulations examining the impact of methyl and ethyl nitrates on tropospheric ozone chemistry in remote regions have been conducted by \citep{Neu2008}. In their study, they incorporated methyl and ethyl nitrates in the chemical mechanism of the UCI CTM [Carver et al., 1997] and by matching their observed atmospheric abundances derived a total oceanic emission flux of these two nitrates (0.35 Tg of N per year). Comparing two simulations with and without spatially uniform emissions of methyl and ethyl nitrates they found that the largest impact of oceanic alkyl nitrates is seen over the equatorial Western Pacific (Figure ???). Atmospheric circulation (convergence of the trade winds) in this region isolates the marine boundary layer, resulting in sustained photochemical loss of $O_3$ (Figure ??? c) and at the same time accumulation of alkyl nitrates (Figure ??? a), which drive up boundary layer $NO_x$ abundances by as much as 250\% (Figure ??? b). Such combination of factors leads to increases in $O_3$ of up to 20\% (Figures ??? c and d), with the largest increases occurring from May to October.

\begin{figure}\label{fig:Neu2008}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT_colorbar}.png}
\end{figure}
\begin{figure}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.2_MeONO2_OCT}.png}
\end{figure}
\begin{figure}
\caption{1}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_NOxdiff}.png}
\end{figure}
\begin{figure}
\caption{October monthly mean mixing ratios of (a) methyl nitrate ($MeONO_2$, $CH_3ONO_2$) and (c) $O_3$, and percent difference (perturbation - control) in (b) $NO_x$, (d) $O_3$. All averaged over the boundary layer (below 850 mb). Adapted from \citep{Neu2008}}
\includegraphics[width=.5\linewidth]{{D:/UEA/ENV-MB4X_Dissertation/ENV-MB4X_Dissertation_thesis/ENV-MB4X_Dissertation_thesis_latex/pics/Neu2008_Fig.3_OCT_O3contol_O3diff}.png}
\end{figure}
